The Use of Quantum Computing\\ in Algorithmic Trading

### 2025-01-05 Documents/Article/article.tex
zpb(article): v1.1

\section{Algorithms Overview}
\subsection{Principal Component Analysis (PCA)}
Principal Component Analysis (PCA) is a widely-used dimensionality reduction technique that allows for the simplification of complex datasets while retaining as much of the variance as possible. The goal of PCA is to transform a dataset into a new coordinate system, where the axes (known as principal components) correspond to the directions of maximum variance in the data. By projecting data onto a lower-dimensional space, PCA reduces the number of features needed to represent the data, making it easier to analyze and visualize. This process is particularly useful in fields such as finance, where large amounts of historical market data may include redundant or highly correlated features.

@@

\subsection{Support Vector Machines (SVM)}
Support Vector Machines (SVMs) are a class of supervised learning algorithms primarily used for classification and regression tasks. The main idea behind SVM is to find the optimal hyperplane that best separates data points belonging to different classes or, in the case of regression, fits the data with minimal error. In its simplest form, SVM constructs a hyperplane in a high-dimensional space that maximizes the margin between two classes. For regression tasks, SVM aims to find a function that has the smallest deviation from the actual data points within a predefined margin of tolerance.

@@

\subsection{Quantum Principal Component Analysis (QPCA)}
Quantum Principal Component Analysis (QPCA) is an extension of classical PCA that leverages quantum mechanics to accelerate the process of dimensionality reduction. The basic principle behind QPCA is to use quantum superposition and entanglement to perform linear algebra operations in parallel, which can significantly reduce the computational time required to analyze large datasets. While classical PCA typically requires matrix diagonalization, a process that grows increasingly complex as the size of the data increases, QPCA uses quantum states to encode and process the data in a way that allows for exponential speedup under certain conditions.

@@

\subsection{Quantum Support Vector Machines (QSVM)}
Quantum Support Vector Machines (QSVM) are an adaptation of classical SVM algorithms designed to take advantage of quantum computing's unique capabilities. While traditional SVM relies on kernel functions to map input data into a higher-dimensional space, QSVM utilizes quantum kernels that are based on quantum states, allowing the model to operate in exponentially larger feature spaces. This enables QSVM to capture more complex, nonlinear relationships in data, which is especially important in the context of stock market prediction, where relationships between features can be highly intricate and non-linear.

@@

\textit{QPCA:} QPCA demonstrates error behavior that is comparable to SVM. While the errors are not as erratic as those of QSVM, they remain significant and follow a similar pattern to SVM, indicating that QPCA also struggles to generalize effectively and adapt to dynamic market conditions.

@@

\begin{table}[ht!]
    \centering
    \begin{tabular}{|l|c|c|c|c|}
        \hline
        \textbf{Metric} & \textbf{PCA} & \textbf{SVM} & \textbf{QPCA} & \textbf{QSVM} \\
        \hline
        Max Abs. Error & 203.26 & 3315.49 & 3399.27 & 4164.38 \\
        Min Abs. Error & -313.06 & 29.24 & 814.45 & 363.91 \\
        Mean Abs. Error & 1.47 & 1803.85 & 2014.48 & 2136.99 \\
        Median Abs. Error & 2.39 & 1794.61 & 1868.78 & 1973.63 \\
        Max Rel. Error & 0.080 & 0.691 & 0.710 & 0.923 \\
        Min Rel. Error & -0.131 & 0.013 & 0.364 & 0.160 \\
        Mean Rel. Error & 0.00035 & 0.489 & 0.567 & 0.595 \\
        Median Rel. Error & 0.00076 & 0.548 & 0.570 & 0.607 \\
        Mean Sqr. Error & 1624.12 & $4.1 \times 10^6$ & $4.6 \times 10^6$ & $5.4 \times 10^6$ \\
        \hline
    \end{tabular}
    \vspace{1em}
    \caption{Error metrics for PCA, SVM, QPCA, and QSVM on the S\&P 500 dataset.}
    \label{tab:error_metrics}
\end{table}

@@

Figure~\ref{fig:pca_trader_portfolio} shows the predicted portfolio value over time for each trader type using the PCA model. The results indicate that the best-performing strategy is the \textbf{Buy and Keep Trader}, as the portfolio steadily increases over time, reflecting the overall upward trend in the market. The 100\% allocation for this trader ensures they capture the full value of the market's growth. Similarly, the \textbf{Remain Value Trader} with a 100\% allocation also sees substantial growth, though the growth rate is slightly lower than that of the Buy and Keep strategy. The 5\% and 20\% variations of this strategy exhibit less dramatic growth, with smaller allocations leading to slower portfolio value accumulation.

@@

@@

>>>>

\section{Algorithms Overview}
\subsection{Principal Component Analysis (PCA)}
Principal Component Analysis (PCA) \cite{pca} is a widely-used dimensionality reduction technique that allows for the simplification of complex datasets while retaining as much of the variance as possible. The goal of PCA is to transform a dataset into a new coordinate system, where the axes (known as principal components) correspond to the directions of maximum variance in the data. By projecting data onto a lower-dimensional space, PCA reduces the number of features needed to represent the data, making it easier to analyze and visualize. This process is particularly useful in fields such as finance, where large amounts of historical market data may include redundant or highly correlated features.

@@

\subsection{Support Vector Machines (SVM)}
Support Vector Machines (SVMs) \cite{svm} are a class of supervised learning algorithms primarily used for classification and regression tasks. The main idea behind SVM is to find the optimal hyperplane that best separates data points belonging to different classes or, in the case of regression, fits the data with minimal error. In its simplest form, SVM constructs a hyperplane in a high-dimensional space that maximizes the margin between two classes. For regression tasks, SVM aims to find a function that has the smallest deviation from the actual data points within a predefined margin of tolerance.

@@

\subsection{Quantum Principal Component Analysis (QPCA)}
Quantum Principal Component Analysis (QPCA) \cite{qpca} is an extension of classical PCA that leverages quantum mechanics to accelerate the process of dimensionality reduction. The basic principle behind QPCA is to use quantum superposition and entanglement to perform linear algebra operations in parallel, which can significantly reduce the computational time required to analyze large datasets. While classical PCA typically requires matrix diagonalization, a process that grows increasingly complex as the size of the data increases, QPCA uses quantum states to encode and process the data in a way that allows for exponential speedup under certain conditions.

@@

\subsection{Quantum Support Vector Machines (QSVM)}
Quantum Support Vector Machines (QSVM) \cite{qsvm} are an adaptation of classical SVM algorithms designed to take advantage of quantum computing's unique capabilities. While traditional SVM relies on kernel functions to map input data into a higher-dimensional space, QSVM utilizes quantum kernels that are based on quantum states, allowing the model to operate in exponentially larger feature spaces. This enables QSVM to capture more complex, nonlinear relationships in data, which is especially important in the context of stock market prediction, where relationships between features can be highly intricate and non-linear.

@@

\textbf{QPCA:} QPCA demonstrates error behavior that is comparable to SVM. While the errors are not as erratic as those of QSVM, they remain significant and follow a similar pattern to SVM, indicating that QPCA also struggles to generalize effectively and adapt to dynamic market conditions.

@@

\begin{table}[ht!]
    \centering
    \begin{tabular}{|l|c|c|c|c|}
        \hline
        \textbf{Metric} & \textbf{PCA} & \textbf{SVM} & \textbf{QPCA} & \textbf{QSVM} \\
        \hline
        Max Abs. Error & 203.26 & 3315.49 & 3399.27 & 4164.38 \\
        Min Abs. Error & -313.06 & 29.24 & 814.45 & 363.91 \\
        Mean Abs. Error & 1.47 & 1803.85 & 2014.48 & 2136.99 \\
        Median Abs. Error & 2.39 & 1794.61 & 1868.78 & 1973.63 \\
        Max Rel. Error & 0.080 & 0.691 & 0.710 & 0.923 \\
        Min Rel. Error & -0.131 & 0.013 & 0.364 & 0.160 \\
        Mean Rel. Error & 0.00035 & 0.489 & 0.567 & 0.595 \\
        Median Rel. Error & 0.00076 & 0.548 & 0.570 & 0.607 \\
        Mean Sqr. Error & 1624.12 & $4.1 \times 10^6$ & $4.6 \times 10^6$ & $5.4 \times 10^6$ \\
        \hline
    \end{tabular}
    \vspace{2.4em}
    \caption{Error metrics for PCA, SVM, QPCA, and QSVM on the S\&P 500 dataset.}
    \label{tab:error_metrics}
\end{table}

@@

Figure~\ref{fig:pca_trader_portfolio} shows the predicted portfolio value over time for each trader type using the PCA model. The results indicate that the best-performing strategy is the Buy and Keep Trader, as the portfolio steadily increases over time, reflecting the overall upward trend in the market. The 100\% allocation for this trader ensures they capture the full value of the market’s growth.

Similarly, the Remain Value Trader with a 100\% allocation also sees substantial growth, though the growth rate is slightly lower than that of the Buy and Keep strategy. The 5\% and 20\% variations of this strategy exhibit less dramatic growth, with smaller allocations leading to slower portfolio value accumulation.

@@

Notably, in the case of individual stocks, such as Nvidia, traders employing the 100\% Remain Value strategy were able to achieve exceptional results, significantly outperforming the index. For example, the portfolio value for Nvidia doubled relative to the overall market trend. This demonstrates that while broader strategies like Buy and Keep capture general market growth, focused approaches on specific high-performing stocks can yield substantial profits.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{pca_nvidia_traders.png}
    \caption{Predicted portfolio value over time for Nvidia stock across different trader types using the PCA model.}
    \label{fig:pca_nvidia_trader_portfolio}
\end{figure}

@@

\bibitem{pca} 
Jolliffe, I. T. (2002). \textit{Principal Component Analysis}. Springer. \url{https://doi.org/10.1007/b98835}

\bibitem{svm} 
Cortes, C., \& Vapnik, V. (1995). Support-vector networks. \textit{Machine Learning}, 20(3), 273–297. \url{https://doi.org/10.1007/BF00994018}

\bibitem{qpca} 
Lloyd, S., Mohseni, M., \& Rebentrost, P. (2014). Quantum principal component analysis. \textit{Nature Physics}, 10, 631–633. \url{https://doi.org/10.1038/nphys3029}

\bibitem{qsvm} 
Rebentrost, P., Mohseni, M., \& Lloyd, S. (2014). Quantum Support Vector Machine for Big Data Classification. \textit{Physical Review Letters}, 113(13), 130503. \url{https://doi.org/10.1103/PhysRevLett.113.130503}
